WHOOMP! (There It Is) \\ Rapid Bayesian position reconstruction for gravitational\nobreakdashes-wave transients

### 2014-07-19 ms.tex
Fix error in weights
Minor edits

In the frequency domain, the strain observed by a single \ac{GW} interferometer is
%
\begin{equation}\label{eq:signal-model}
    x_i (\omega) = h_i (\omega; \boldsymbol\theta) + n_i (\omega),
\end{equation}
%
where $h_i (\omega; \boldsymbol\theta)$ is the \ac{GW} signal given a parameter vector $\boldsymbol\theta$ that describes the \ac{GW} source, and $n_i (\omega)$ is that detector's \ac{WSS} Gaussian noise with one\nobreakdashes-sided \ac{PSD} $S_i(\omega) = E\left[\left|n_i(\omega)\right|^2\right]$. We shall denote the combined observation from a network of detectors as $\mathbf x (\omega) \equiv \{x_i (\omega)\}_i$.

@@

Assuming circular orbits and no spin precession, we can write the \ac{GW} signal in each detector as a linear combination of two basis waveforms $h_0$ = $h$ and $h_{\pi/2}$. For nonprecessing systems, $h_0$ and $h_{\pi/2}$ are approximately ``in quadrature'' in the same sense as the sine and cosine functions, being nearly orthogonal and out of phase by ${\pi/2}$ at all frequencies. If $h_0$ and $h_{\pi/2}$ are Fourier transforms of real functions, then $h_0(\omega) = h_0^*(-\omega)$ and $h_{\pi/2}(\omega) = h_{\pi/2}^*(-\omega)$, and we can write (assuming an arbitrary phase convention)
%
\begin{equation}
    h_{\pi/2}(\omega) = h_0(\omega) \cdot
    \cases{
        -i & if $\omega \geq 0$ \\
        i & if $\omega < 0$
    }.
\end{equation}
%
Then, we can write the signal model in a way that isolates all dependence on the extrinsic parameters $\boldsymbol\theta_\mathrm{ex}$ into the coefficients and all dependence on the intrinsic parameters $\boldsymbol\theta_\mathrm{in}$ into the basis waveform, by taking the Fourier transform of Equation~(2.8) of \cite{PhysRevD.83.084002}:
%
\begin{equation}\label{eq:full-signal-model}
    h_i(\omega; \boldsymbol\theta) = \e^{-i \omega t_\oplus}
    \frac{D_{1,i}}{D_\mathrm{L}}
    \e^{2 i \phi_c}
    \left[
    \frac{1}{2} \left(1 + \cos^2 \iota\right) \Re \left\{z\right\} - i
    \left(\cos\iota\right) \Im z
    \right]
    h(\omega; \boldsymbol\theta_\mathrm{in})
\end{equation}
%
for $\omega \geq 0$, where
%
\begin{equation}
    z = e^{-2 i \psi} \left(
    F_{+,i}(\alpha, \delta, t_\oplus) +
    i F_{\times,i}(\alpha, \delta, t_\oplus)
    \right).
\end{equation}
%
The quantities $F_{+,i}$ and $F_{\times,i}$ are the dimensionless detector antenna factors, defined such that $0 \leq {F_{+,i}}^2 + {F_{\times,i}} \leq 1$. They depend on the orientation of detector $i$ as well as the sky location and sidereal time of the event, and are derived in~\cite{PhysRevD.63.042003}. The quantity $D_{1,i}$ is a fiducial distance at which detector $i$ would register SNR=1 for an optimally oriented binary (face\nobreakdashes-on, and in a direction perpendicular to the interferometer's arms),
%
\marginpar{I seem to have dropped a factor of 2 here...}%
%
\begin{equation}\label{eq:horizon}
D_{1,i} = 1 / \sigma_i, \qquad {\sigma_i}^2 = \int_0^\infty \frac{\left|h(\omega; \mathbf \theta)\right|^2}{S_i(\omega)} \,d\omega.
\end{equation}

@@

\begin{figure}
    \begin{center}
        \includegraphics{crlb_tau}
    \end{center}
    \caption{\label{fig:crlb-tau}\ac{CRLB} on root\nobreakdashes-mean square timing uncertainty, using the likelihood for the full GW data (Equation~\ref{eq:gaussian-likelihood-spa}; dashed diagonal line) or the autocorrelation likelihood (Equation~\ref{eq:autocor-likelihood}; solid lines) with a selection of arrival time priors.}
\end{figure}

@@

The arrival time prior is connected to the origin of the detector coordinate system. Given the Earth-fixed coordinates of the detectors $\mathbf{n}_i$ and the arrival times $\tau_i$, we compute their averages weighted by the timing uncertainty formula:
%
\begin{equation*}
    \langle \mathbf{n} \rangle = \frac{
        \displaystyle
        \sum_i \frac{\mathbf{n}_i}{\hat\rho_i \omega_{\mathrm{rms},i}}
    }{
        \displaystyle
        \sum_i \frac{1}{\hat\rho_i \omega_{\mathrm{rms},i}}
    },
    \qquad
    \langle \hat\tau \rangle = \frac{
        \displaystyle
        \sum_i \frac{\hat\tau_i}{\hat\rho_i \omega_{\mathrm{rms},i}}
    }{
        \displaystyle
        \sum_i \frac{1}{\hat\rho_i \omega_{\mathrm{rms},i}}
    }.
\end{equation*}
%
Then, we subtract off these means:
%
\begin{equation*}
    \mathbf{n}_i \leftarrow \mathbf{n}_i - \langle \mathbf{n} \rangle,
    \qquad
    \hat\tau_i \leftarrow \hat\tau_i - \langle \hat\tau \rangle.
\end{equation*}
%
In these coordinates relative to the weighted detector array barycenter, the arrival time prior is uniform in $-T \leq t \leq T$, with $T = \max\limits_i |\mathbf{n}_i| / c + 5\,\textrm{ms}$.

>>>>

In the frequency domain, the strain observed by a single \ac{GW} interferometer is
%
\begin{equation}\label{eq:signal-model}
    x_i (\omega) = h_i (\omega; \boldsymbol\theta) + n_i (\omega),
\end{equation}
%
where $h_i (\omega; \boldsymbol\theta)$ is the \ac{GW} signal given a parameter vector $\boldsymbol\theta$ that describes the \ac{GW} source, and $n_i (\omega)$ is that detector's \ac{WSS} Gaussian noise with one\nobreakdashes-sided \ac{PSD} $S_i(\omega) = E\left[\left|n_i(\omega)\right|^2\right] + E\left[\left|n_i(-\omega)\right|^2\right] = 2 E\left[\left|n_i(\omega)\right|^2\right]$. We shall denote the combined observation from a network of detectors as $\mathbf x (\omega) \equiv \{x_i (\omega)\}_i$.

@@

Assuming circular orbits and no spin precession, we can write the \ac{GW} signal in each detector as a linear combination of two basis waveforms $h_0$ and $h_{\pi/2}$. For nonprecessing systems, $h_0$ and $h_{\pi/2}$ are approximately ``in quadrature'' in the same sense as the sine and cosine functions, being nearly orthogonal and out of phase by ${\pi/2}$ at all frequencies. If $h_0$ and $h_{\pi/2}$ are Fourier transforms of real functions, then $h_0(\omega) = h_0^*(-\omega)$ and $h_{\pi/2}(\omega) = h_{\pi/2}^*(-\omega)$, and we can write (assuming an arbitrary phase convention)
%
\begin{equation}
    h_{\pi/2}(\omega) = h_0(\omega) \cdot
    \cases{
        -i & if $\omega \geq 0$ \\
        i & if $\omega < 0$
    }.
\end{equation}
%
Then, we can write the signal model in a way that isolates all dependence on the extrinsic parameters $\boldsymbol\theta_\mathrm{ex}$ into the coefficients and all dependence on the intrinsic parameters $\boldsymbol\theta_\mathrm{in}$ into the basis waveform, by taking the Fourier transform of Equation~(2.8) of \cite{PhysRevD.83.084002}:
%
\begin{equation}\label{eq:full-signal-model}
    h_i(\omega; \boldsymbol\theta) = \e^{-i \omega t_\oplus}
    \frac{D_{1,i}}{D_\mathrm{L}}
    \e^{2 i \phi_c}
    \left[
    \frac{1}{2} \left(1 + \cos^2 \iota\right) \Re \left\{z\right\} - i
    \left(\cos\iota\right) \Im z
    \right]
    h(\omega; \boldsymbol\theta_\mathrm{in})
\end{equation}
%
for $\omega \geq 0$, where
%
\begin{equation}
    z = e^{-2 i \psi} \left(
    F_{+,i}(\alpha, \delta, t_\oplus) +
    i F_{\times,i}(\alpha, \delta, t_\oplus)
    \right).
\end{equation}
%
The quantities $F_{+,i}$ and $F_{\times,i}$ are the dimensionless detector antenna factors, defined such that $0 \leq {F_{+,i}}^2 + {F_{\times,i}} \leq 1$. They depend on the orientation of detector $i$ as well as the sky location and sidereal time of the event, and are derived in~\cite{PhysRevD.63.042003}. The quantity $D_{1,i}$ is a fiducial distance at which detector $i$ would register SNR=1 for an optimally oriented binary (face\nobreakdashes-on, and in a direction perpendicular to the interferometer's arms),
%
\marginpar{I seem to have dropped a factor of 2 here...}%
%
\begin{equation}\label{eq:horizon}
D_{1,i} = 1 / \sigma_i, \qquad {\sigma_i}^2 = \int_0^\infty \frac{\left|h(\omega; \mathbf \theta)\right|^2}{S_i(\omega)} \,d\omega.
\end{equation}

@@

\begin{figure}
    \begin{center}
        \includegraphics{crlb_tau}

        \includegraphics{crlb_gamma}
    \end{center}
    \caption{\label{fig:crlb-tau}\ac{CRLB} on root\nobreakdashes-mean square timing uncertainty, using the likelihood for the full GW data (Equation~\ref{eq:gaussian-likelihood-spa}; dashed diagonal line) or the autocorrelation likelihood (Equation~\ref{eq:autocor-likelihood}; solid lines) with a selection of arrival time priors.}
\end{figure}

@@

The arrival time prior is connected to the origin of the detector coordinate system. Given the Earth-fixed coordinates of the detectors $\mathbf{n}_i$ and the arrival times $\tau_i$, we compute their averages weighted by the timing uncertainty formula:
%
\begin{equation*}
    \langle \mathbf{n} \rangle = \frac{
        \displaystyle
        \sum_i \frac{\mathbf{n}_i}
            {\left(\hat\rho_i \omega_{\mathrm{rms},i}\right)^2}
    }{
        \displaystyle
        \sum_i \frac{1}{\left(\hat\rho_i \omega_{\mathrm{rms},i}\right)^2}
    },
    \qquad
    \langle \hat\tau \rangle = \frac{
        \displaystyle
        \sum_i \frac{\hat\tau_i}
            {\left(\hat\rho_i \omega_{\mathrm{rms},i}\right)^2}
    }{
        \displaystyle
        \sum_i \frac{1}{\left(\hat\rho_i \omega_{\mathrm{rms},i}\right)^2}
    }.
\end{equation*}
%
Then, we subtract off these means:
%
\begin{equation*}
    \mathbf{n}_i \leftarrow \mathbf{n}_i - \langle \mathbf{n} \rangle,
    \qquad
    \hat\tau_i \leftarrow \hat\tau_i - \langle \hat\tau \rangle.
\end{equation*}
%
In these coordinates relative to the weighted detector array barycenter, the arrival time prior is uniform in $-T \leq t \leq T$, with $T = \max\limits_i |\mathbf{n}_i| / c + 5\,\textrm{ms}$.
